Probability Aggregation: Dynamic Hierarchical Modeling of Expert Beliefs in Geopolitical Events

### 2013-08-23 paper.tex
Edits

\begin{abstract}
Most subjective probability aggregation procedures use a single
probability judgement from each expert, even though it is common for
experts studying real problems to update their probability estimates
over time.  This paper makes an advance towards unexplored areas of
probability aggregation by considering a dynamic context in which
experts are allowed to update their beliefs at random intervals.  We
demonstrate our method on a real-world dataset that includes over
2,300 experts making multiple probability forecasts on different
subsets of 166 different international political events. We present a
dynamic model that takes into account the forecaster’s level of
self-reported expertise and produces aggregate probabilities that are
sharp and well-calibrated both in- and out-of-sample. Using the
aggregate probabilities as the no-bias reference-line, the lower the
self-reported expertise in a group, the more under-confident it is
found to be.
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of their assessments is typically evaluated with an important attribute called \textit{calibration}. An expert is said to be well-calibrated if 
the proportion of events that actually occurred is, say, 60\% for all those events for which he assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), some personal differences among the experts have been found to associate with calibration. In particular, \citet{wright1994coherence} argue that a higher self-rated expertise on the subject matter implies better calibration.

Unfortunately, calibration by itself does not guarantee usefulness of a probability assessment. For instance, if the empirical frequency of  the last 10 years of a relatively stationary process is 45\%, the expert can assess a constant probability of 0.45 and turn out well-calibrated. His probability, however, is not very useful in decision making. In general, a probability assessments that is near 0.0 or 1.0, is considered more informative, and an expert who makes such assessments is called \textit{sharp}. In the forecasting literature sharpness is often combined with calibration into a single goal of maximizing sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}). An expert that meets this goal is considered to give perfect probability forecasts (\citet{gneiting2004verifying}). Even though expecting perfect forecasts from an expert is not reasonable in practice, this criterion has led to a wide range of novel and insightful observations in probability forecasting.

One such observation is related to the aggregation of multiple probabilities. There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  results in better predictive performance. Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the experts' advice only once before the future event resolves.

Therefore many complex areas of probability aggregation still remain rather unexplored. For instance, consider an investor aiming to assess whether a stock index will finish trading at or above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts every day and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments on a daily basis, the aggregation should be performed by taking into account the correlation and time-series nature of their advice. Analyzing this kind of future events under daily updating of the forecasts is the main focus of this paper. The core research goal is to construct an interpretable time-series model that incorporates self-reported expertise in order to estimate a sharp and well-calibrated version of the crowd belief. The model is then used for
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a number of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure: First, constrained parameter estimates are obtained via Gibbs sampling. These estimates are then transformed in to their unconstrained equivalents by a calibrating step. An extension of this model to polychotomous outcomes is briefly discussed before the model evaluation sections. The first evaluation section uses synthetic data to study the accuracy of the two-step procedure to capture parameter values. The second and much longer evaluation section uses the model on our probability forecasting data. The main goal is to illustrate and also to study the specific research objectives itemized above. The paper concludes with a discussion on future research directions and model limitations.

\section{Geopolitical Forecasting Data}
\label{data}
The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to forecasting and political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points are positively associated with the self-reported expertise levels. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active. Updating, however, was done on a very infrequent basis. An expert gave, on average, only 0.0159 forecasts per day. A typical expert, therefore, made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

@@

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Problem is Active  & 4.00 &  35.25 &  72.00 & 106.30 & 145.20 & 418.00 \\
\# of Experts per Problem & 212.0 &   543.2 &  693.5 &  783.7 &  983.2 & 1690.0\\ 
\# Forecasts given by each Expert on a Problem & 1.0 &  1.0 &   1.0  &  1.8 &  2.0 & 131.0 \\
\# Problems participated by an Expert & 1.0 &  14.0 &   36.0  &  55.0 &  90.0 & 166.0 \\ \hline
\end{tabular}
\caption{Five-number summaries of the real-world data used in this paper: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.}
\label{DataStats}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\caption{Frequencies of the expertise levels across all the 166 questions in the dataset.}
\label{ExpertiseTable}
\end{center}
\end{table}

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with $Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R$ and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$, the error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen a priori. The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the forecasters are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. The bias terms are analyzed and discussed more in Section \ref{ExpertBias}. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible, yet efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable and the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via a Gibbs sampler that only makes use of standard distributions. This procedure is standard practice in Bayesian statistics and is therefore not described in great detail here. See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.

>>>>

\begin{abstract}
Most subjective probability aggregation procedures use a single probability judgement from each expert, even though it is common for experts studying real problems to update their probability estimates over time.  This paper makes an advance towards unexplored areas of probability aggregation by considering a dynamic context in which experts are allowed to update their beliefs at random intervals. The updates are assumed to occur very infrequently, resulting into a highly sparse dataset that cannot be modeled by standard time-series procedures. In response to the lack of appropriate methodology, this paper presents a dynamic model that takes into account the forecaster’s level of self-reported expertise and produces aggregate probabilities that are sharp and well-calibrated both in- and out-of-sample. The model is demonstrated on a real-world dataset that includes over 2,300 experts making multiple probability forecasts on different subsets of 166 different international political events. 
\end{abstract}

@@

\section{Introduction}
\label{intro}
Individual experts can differ radically from one another in their abilities to assess probabilities of future events. The quality of a probability assessment is typically evaluated in regards to an important attribute called \textit{calibration}. An expert is said to be well-calibrated if his probabilities agree with the long-term frequency of the events. This means that, for instance, the proportion of events that actually occurred is 60\% for all those events for which he assessed a probability of 0.60. Even though several experiments have been conducted to show that experts are generally poorly calibrated (see, e.g., \citet{cooke1991experts, shlyakhter1994quantifying}), the level of expertise has been found to be associated with calibration. In particular, \citet{wright1994coherence} argue that a higher self-reported expertise on the subject matter implies better calibration.

Unfortunately, calibration by itself does not guarantee usefulness. To see this, consider a relatively stationary process, such as rain in some geographic region, whose empirical frequency of occurrence in the last 10 years is 45\%. In this setting, an expert can assess a constant probability of 0.45 and be well-calibrated. This probability, however, is very near 0.5 and hence too uninformative to be useful in decision making. Compare this to the advice of  another expert whose typical probability assessment is near 0.0 or 1.0; that is, he is a \textit{sharp} assessor. If the expert is both sharp and well-calibrated, then he is able to forecast the behavior of the process both with high certainty and accuracy. From the decision-maker's perspective, such advice is highly useful as it involves very little risk. Therefore the goal in probability estimation should be to maximize sharpness subject to calibration (see, e.g., \citet{raftery2005using, Ranjan08}). This is a well-known goal that has led to a wide range of novel and insightful observations in probability forecasting literature.

One such observation is related to the aggregation of multiple probabilities. There is strong empirical evidence that bringing together the strengths of different experts by combining their probability forecasts into a single consensus, known as the \textit{crowd belief},  results in better predictive performance. Being motivated by the long list of applications of probability forecasts, including medical diagnosis (\citet{wilson1998prediction, pepe2003statistical}, political and socio-economic foresight (\citet{tetlock2005expert}), and meteorology (\citet{sanders1963subjective, vislocky1995improved, baars2005performance}), researchers have proposed many different approaches to combining probability forecasts (see, e.g., \citet{Ranjan08, satopaa} for some recent studies, and \citet{Genest, Wallsten97evaluatingand, clemen2007aggregating, primo2009calibration} for a comprehensive overview). The general focus, however, has been on developing one-time aggregation procedures that consult the expert's advice only once before the event resolves.

Consequently, many areas of probability aggregation still remain rather unexplored. For instance, consider an investor aiming to assess whether a stock index will finish trading above some threshold on a given date. In order to maximize his overall predictive accuracy, he may consult a group of experts repeatedly over a period of time and adjust his estimate of the aggregate probability accordingly. Since the experts are allowed to update their probability assessments, the aggregation should be performed by taking into account the temporal correlation in their advice. Many standard time-series procedures could be used to perform this aggregation as long as the experts update their advice consistently on a daily basis.

%If the consulted experts diligently make probability assessments on a daily basis, many previously-studied time-series procedures could be used to aggregate their assessments.

This paper adds another layer of complexity by assuming a heterogeneous set of experts, most of whom only make one or two probability assessments over the hundred or so days before the event resolves. The problem at hand is therefore strikingly different from many time-series estimation problems, where one has an observation at every time point -- or almost every time point.
Therefore standard time-series procedures like ARIMA (see, e.g., \citet{mills1991time}) are not applicable without ignoring the heterogeneity among the experts. In response to the lack of procedures to model sparse and heterogeneous probability forecasting data, this paper introduces an interpretable time-series model that incorporates self-reported expertise and captures a sharp and well-calibrated estimate of the crowd belief. The model is then used for
\begin{itemize}
\item the analysis of group-level under- and overconfidence across different levels of self-reported expertise,
%\item confidence bands for the crowd belief,
\item accurate probability forecasts, and
\item many question-specific quantities that have easy interpretations and can be used to gain novel insight in the social sciences. 
\end{itemize}

The paper begins with a description of our probability forecasting data that were collected by asking over 2,300 experts to give probability forecasts and to self-assess their level of expertise on a subset of 166 geopolitical binary events. After summarizing the dataset, the paper introduces a dynamic hierarchical model for capturing the crowd belief. The model is estimated in a two-step procedure. First, constrained parameter estimates are obtained via Gibbs sampling (see \citet{geman1984stochastic} for the original introduction of Gibbs sampling). These estimates are then transformed to their unconstrained equivalents by a calibration step. An extension of this model to polychotomous outcomes is briefly discussed before model evaluation. The first evaluation section uses synthetic data to study the accuracy to which the two-step procedure is able to estimate parameter values. The second evaluation section discusses different aspects of the model by applying it to our real-world probability forecasting data. The paper concludes with a discussion on future research directions and model limitations.

\section{Geopolitical Forecasting Data}
\label{data}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Statistic & Min. & $Q_1$ & Median & Mean & $Q_3$ & Max.\\ \hline
\# of Days a Problem is Active  & 4 &  35.6 &  72.0 & 106.3 & 145.20 & 418 \\
\# of Experts per Problem & 212 &   543.2 &  693.5 &  783.7 &  983.2 & 1690\\ 
\# Forecasts given by each Expert on a Problem & 1 &  1.0 &   1.0  &  1.8 &  2.0 & 131 \\
\# Problems participated by an Expert & 1 &  14.0 &   36.0  &  55.0 &  90.0 & 166 \\ \hline
\end{tabular}
\caption{Five-number summaries of our real-world data.
%: \textit{\# of Days Active} represents the number of days that the experts were given to update their forecast before the correct answer was revealed; \textit{\# of Experts per Problem } denotes the number of experts participating in a question; \textit{\# Forecasts given by each Expert on a Problem} gives the number of forecasts for a question made by an active expert; \textit{\# Problems participated by an Expert} represents the number of different questions a given expert participated in.
}
\label{DataStats}
\end{center}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c c }
\hline
Expertise Level & 1 & 2 & 3 & 4 & 5\\
Frequency (\%) & 25.3 & 30.7 & 33.6 &  8.2 &  2.1 \\\hline
\end{tabular}
\caption{Frequencies of the self-reported expertise (1 = Not At All Expert and 5 = Extremely Expert) levels across all the 166 questions in our real-world data.}
\label{ExpertiseTable}
\end{center}
\end{table}

The data collection began with a recruitment of 2,365 forecasters ranging from graduate students to political science faculty and practitioners. The recruiting was made from professional societies, research centers, alumni associations, science bloggers, and word of mouth. Requirements included at least a Bachelor's degree and completion of psychological and political tests that took roughly two hours. These measures assessed cognitive styles, cognitive abilities, personality traits, political attitudes, and real-world knowledge. The experts were asked to give probability forecasts (to the second decimal point) and to self-assess their level of expertise (measured on a 1-to-5 scale with 1 = Not At All Expert and 5 = Extremely Expert) on a number of 166 geopolitical binary events taking place between September 29, 2011 and May 8, 2013. Each question was active for a period of time during which the participating experts were allowed to update their forecasts as frequently as they liked without being penalized. The forecasters knew that their probability estimates would be assessed for accuracy using Brier scores\footnote{The Brier score is the squared distance between the probability forecast and the event indicator that equals 1.0 or 0.0 depending on whether the event happened or not, respectively. See \citet{Brier} for the original introduction.}. This incentivized them to report their true beliefs instead of attempting to game the system (\citet{winkler1968good}). In addition to receiving \$150 for meeting minimum participation requirements that did not depend on prediction accuracy, the forecasters also received status rewards for their performance via leader-boards displaying Brier scores for the top 20 forecasters. Since a typical expert participated only in a small subset of the 166 questions, the forecasters are considered indistinguishable conditional on the self-reported level of expertise.
Tables \ref{DataStats} and \ref{ExpertiseTable} provide more relevant summary statistics on the data. Notice that the distribution of the self-reported expertise is skewed to the right and that some questions remained active longer than others.  For more details on the data and its collection see \citet{ungar2012good}.

@@

To illustrate the nature of the data with some concrete examples, Figures \ref{Example1} and \ref{Example2} show scatterplots of the probability forecasts given for (a) \textit{Will the expansion of the European bailout fund be ratified by all 17 Eurozone nations before 1 November 2011?}, and (b) \textit{Will the Nikkei 225 index finish trading at or above 9,500 on 30 September 2011?}. The points have been jittered slightly to make overlaps visible. The darkness of the points are positively associated with the self-reported expertise levels. Given that the European bailout fund was ratified before 1 November 2011 and that the Nikkei 225 index finished trading at around 8,700 on 30 September 2011, the general trend of the probability forecasts converge to the correct answers. There is, however, much disagreement among the experts. This disagreement persists even near the closing dates of the questions. Unfortunately, the data is not only noisy but also quite sparse. As mentioned before, the forecasters were allowed to update their forecast as long as the questions remained active. Updating was done on a very infrequent basis. An expert gave, on average, only 0.0159 forecasts per day, and so a typical expert made a forecast about every 62 days. As a result the average response rate was around 13.5 forecasts per day from a large group of experts.

\section{Model}
\label{model}
Let $p_{i,t,k} \in (0,1)$ be the probability forecast given by the $i$th expert at time $t$ for the $k$th question, where $i = i, \dots, I_k$, $t = 1, \dots, T_k$, and $k = 1, \dots, K$. Denote the logit-probabilities with 
\begin{eqnarray*}
Y_{i,t,k} = \logit(p_{i,t,k}) = \log \left( \frac{p_{i,t,k}}{1-p_{i,t,k}}\right) \in \R
\end{eqnarray*}
and collect the logit-probability forecasts given for question $k$ at time $t$ into a vector $\boldsymbol{Y}_{t,k} = [Y_{1,t,k}\ Y_{2,t,k}\ \dots\ Y_{I_{k},t,k}]^T$. Partition the experts into $J$ groups based on some individual feature, such as self-reported expertise, with each group sharing a common multiplicative bias term $b_{j} \in \R$ for $j = 1, \dots, J$. Collect these bias terms into a bias vector $\boldsymbol{b} = [b_{1}\ b_{2}\ \dots\ b_{J}]^T$ and let $\boldsymbol{M}_k$ be a $I_k \times J$ matrix denoting  the group-memberships of the forecasters in question $k$. That is, if the $i$th expert participating in the $k$th question belongs to the $j$th group, then the $i$th row of $\boldsymbol{M}_k$ is the $j$th standard basis vector, $\boldsymbol{e}_j$. The bias vector $\boldsymbol{b}$ does not include a subindex because it is considered shared among all the $K$ questions. To secure model identifiability, it is sufficient to share only one of the elements of $\boldsymbol{b}$ among the questions. This element defines a baseline under which it is possible to estimate the remaining $J-1$ bias terms separately within each of the questions. In this paper, however, the entire vector $\boldsymbol{b}$ is shared because some of the questions in our real-world data set involve very few experts with the highest level of self-reported expertise. Under this notation, the model for the $k$th question can be expressed as 
\begin{eqnarray}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} X_{t, k} + \boldsymbol{v}_{t, k} \label{observedpr} \\
X_{t, k} &=& \gamma_k X_{t-1, k} + w_{t, k} \label{hiddenpr}\\
X_{0,k} &\sim& \mathcal{N}(\mu_0, \sigma^2_0) \nonumber
\end{eqnarray}
where Equation (\ref{observedpr}) denotes the observed process, Equation (\ref{hiddenpr}) shows the hidden process that is driven by the constant $\gamma_k \in \R$. The error terms are independent and identically distributed normal random variables with mean zero
\begin{eqnarray*}
\boldsymbol{v}_{t, k} | \sigma^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k &\stackrel{i.i.d.}{\sim}& \mathcal{N}(0, \tau^2_k),
\end{eqnarray*}
and $(\mu_0, \sigma_0^2) \in (\R, \R^+)$ are hyper-parameters chosen \textit{a priori}. The hidden state $X_{t,k}$ represents the sharp and calibrated logit-probability for the $k$th event given the information available up to and including time $t$. To make this more specific, let $Z_k \in \{0, 1\}$ indicate whether the event associated with the $k$th question happened $\left(Z_k = 1\right)$ or did not happen $\left(Z_k = 0\right)$. If $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ is a filtration representing the information available up to and including a given time point, then $\E[Z_k | \mathcal{F}_{t,k}] = \P(Z_k = 1| \mathcal{F}_{t,k}) = \logit^{-1}(X_{t,k})$. These probabilities are made estimable by assuming that the observed process is associated with the filtration $\{\mathcal{F}_{t,k}\}_{t=1}^{T_k}$ and hence can be used as a proxy for the amount of information available at any given time point. Therefore it is not unreasonable to interpreted the hidden process as the crowd belief across time.

To give intuitive interpretations of the other model parameters, notice that the error term in the observed process has mean zero. Therefore the forecasters are assumed to be, on average, a multiplicative constant $\boldsymbol{b}$ away from the calibrated crowd belief. An individual element of $\boldsymbol{b}$ can be interpreted as a group-specific \textit{systematic bias} that labels the group either as over-confident $\left( b_j \in (1, \infty)\right)$ or as under-confident $\left(b_j \in (0,1)\right)$. Unfortunately, due to the high sparsity of our data, estimating a bias term and analyzing the bias separately for each expert is not possible. See Section \ref{ExpertBias} for an analysis and discussion on the bias terms. Any other deviation from the calibrated crowd belief is considered \textit{random noise}. This noise is measured in terms of $\sigma^2_k$ and can be assumed to be caused by momentary over-optimism (or pessimism), false beliefs, or other misconceptions.

@@

The most challenging part of our model estimation is to incorporate the event indicators $\{Z_k\}_{k=1}^K$ such that the hidden process is well-calibrated and the interpretability of our model is not compromised. This section introduces a two-step procedure, called \textit{Sample-Then-Calibrate} or simply \textit{STC}, that achieves this goal in a flexible and efficient manner by first estimating  the model parameters under a constraint (\textit{Sampling Step}) and then performing a one-dimension optimization procedure to transform the constrained estimates into their unconstrained counterparts (\textit{Calibration Step}).

\subsection{Sampling Step}
\label{sampling_step}
Since $\left(a \boldsymbol{b}, X_{t,k}/a, a^2 \tau_k^2\right) \neq \left(\boldsymbol{b}, X_{t,k}, \tau_k^2\right)$ for any $a > 0$ yield the same  likelihood for $\boldsymbol{Y}_{t,k}$, the model as described by Equations (\ref{observedpr}) and (\ref{hiddenpr}) is not identifiable and the parameter estimates tend to drift during the sampling process. A well-known solution is to choose one of the elements of $\boldsymbol{b}$, say $b_3$, as the reference point and fix $b_3 = 1$.  Denote the constrained version of the model by
\begin{eqnarray*}
\boldsymbol{Y}_{t, k} &=&  \boldsymbol{M}_k \boldsymbol{b} (1) X_{t, k}(1)+ \boldsymbol{v}_{t, k}\\
X_{t, k}(1) &=& \gamma_k(1) X_{t-1, k}(1) + w_{t, k}\\
\boldsymbol{v}_{t, k} | \sigma^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}_{I_k}(\boldsymbol{0}, \sigma^2_k(1) \boldsymbol{I}_{I_k})\\
w_{t, k} | \tau^2_k(1) &\stackrel{i.i.d.}{\sim}& \mathcal{N}\left(0, \tau^2_k(1)\right),
\end{eqnarray*}
where the trailing input parameter emphasizes the constraint $b_{3} = 1$. Since this version is identifiable, estimates of the model parameters can be obtained. Denote the estimates by placing a hat on the parameter symbol. For instance, $\hat{\boldsymbol{b}}(1)$ and $\hat{X}_{t, k}(1)$ represent the estimates of $\boldsymbol{b}(1)$ and $X_{t, k}(1)$, respectively. These estimates are found via a Gibbs sampler that only makes use of standard distributions. See Appendix \ref{appendix} for the technical details of our sampler, and, e.g., \citet{gelman2003bayesian} for a discussion on the general principles of Gibbs sampling.
